---
title: R Plot Gallery
author: R package build
date: '2021-10-24'
slug: test-hide-code
categories: []
tags: []
subtitle: ''
summary: 'In this post, I have listed plenty of plots generated by R. I hope you enjoy this gallery!'
authors: []
lastmod: '2021-10-24T17:35:12+01:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
# output:
#   html_document:
#     theme: flatly
#     highlight: tango
---
<script>
function myFunction(x) {
  if (x.style.display === "none") {
    x.style.display = "block";
  } else {
    x.style.display = "none";
  }
}
</script>
<style>
button{
    background-color: white;
    color: #f44336;
    border-radius:12px;
    outline: 0;
    transition-duration: 0.4s;
    border-color:#f44336;
}

button:hover{
    background-color: #f44336;
    color: white;
}
</style>
```{r echo=FALSE, message=FALSE, results='hide', warning=FALSE}
# import data

library(tidyverse)
library(tidymodels)
library(readr)
library(scales)
library(janitor)
library(tidytuesdayR)
library(cytools)
theme_set(theme_light())

pumpkins_raw <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-19/pumpkins.csv")

water_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-04/water.csv")

pumpkins <-
  pumpkins_raw %>%
  separate(id, into = c("year", "type")) %>%
  mutate(across(c(year, weight_lbs, ott, place), parse_number)) %>%
  filter(type == "P") %>%
  select(weight_lbs, year, place, ott, gpc_site, country)
```

**Note:** Don't be fooled by the subtitle that tells you this article is a 1 min read. 

In this post, I have listed plenty of plots generated by R. You can enjoy these plots as much as you like. If you are interested in how to reproduce these plots, please click on the `Code` button below the plots. If you are interested in what these plots can tell you, please click on the `Explanation` button next to `Code`. I hope you enjoy this gallery!

<figure style="padding: 0rem; margin: 0rem 0rem">
```{r echo=FALSE}
pumpkins %>%
  filter(ott > 20, ott < 1e3) %>%
  ggplot(aes(ott, weight_lbs, color = place)) +
  geom_point(alpha = 0.2, size = 1.1) +
  labs(x = "over-the-top inches", y = "weight (lbs)") +
  scale_color_viridis_c()
```
<button onclick="myFunction(myDIV1)">Code</button>
<div id="myDIV1" style="display: none">
```{r eval=FALSE}
pumpkins %>%
  filter(ott > 20, ott < 1e3) %>%
  ggplot(aes(ott, weight_lbs, color = place)) +
  geom_point(alpha = 0.2, size = 1.1) +
  labs(x = "over-the-top inches", y = "weight (lbs)") +
  scale_color_viridis_c()
```
</div>
</figure> 

<figure style="padding: 0rem; margin: 0rem 0rem">
```{r echo=FALSE}
pumpkins %>%
  filter(ott > 20, ott < 1e3) %>%
  ggplot(aes(ott, weight_lbs)) +
  geom_point(alpha = 0.2, size = 1.1, color = "gray60") +
  geom_smooth(aes(color = factor(year)),
    method = lm, formula = y ~ splines::bs(x, 3),
    se = FALSE, size = 1.5, alpha = 0.6
  ) +
  labs(x = "over-the-top inches", y = "weight (lbs)", color = NULL) +
  scale_color_viridis_d()
```
<button onclick="myFunction(myDIV2)">Code</button>
<div id="myDIV2" style="display: none">
```{r eval=FALSE}
pumpkins %>%
  filter(ott > 20, ott < 1e3) %>%
  ggplot(aes(ott, weight_lbs)) +
  geom_point(alpha = 0.2, size = 1.1, color = "gray60") +
  geom_smooth(aes(color = factor(year)),
    method = lm, formula = y ~ splines::bs(x, 3),
    se = FALSE, size = 1.5, alpha = 0.6
  ) +
  labs(x = "over-the-top inches", y = "weight (lbs)", color = NULL) +
  scale_color_viridis_d()
```
</div>
</figure> 

<figure style="padding: 0rem; margin: 0rem 0rem">
```{r echo=FALSE}
pumpkins %>%
  mutate(
    country = fct_lump(country, n = 10),
    country = fct_reorder(country, weight_lbs)
  ) %>%
  ggplot(aes(country, weight_lbs, color = country)) +
  geom_boxplot(outlier.colour = NA) +
  geom_jitter(alpha = 0.1, width = 0.15) +
  labs(x = NULL, y = "weight (lbs)") +
  theme(legend.position = "none")
```
<button onclick="myFunction(myDIV3)">Code</button>
<div id="myDIV3" style="display: none">
```{r eval=FALSE}
pumpkins %>%
  mutate(
    country = fct_lump(country, n = 10),
    country = fct_reorder(country, weight_lbs)
  ) %>%
  ggplot(aes(country, weight_lbs, color = country)) +
  geom_boxplot(outlier.colour = NA) +
  geom_jitter(alpha = 0.1, width = 0.15) +
  labs(x = NULL, y = "weight (lbs)") +
  theme(legend.position = "none")
```
</div>
<button onclick="myFunction(explain3)">Explanation</button>
<div id="explain3" style="display: none">
dafsafdafdafda
</div>
</figure> 

<figure style="padding: 0rem; margin: 0rem 0rem">
```{r echo=FALSE}
water_raw %>%
  filter(
    country_name == "Sierra Leone",
    lat_deg > 0, lat_deg < 15, lon_deg < 0,
    status_id %in% c("y", "n")
  ) %>%
  ggplot(aes(lon_deg, lat_deg, color = status_id)) +
  geom_point(alpha = 0.1) +
  coord_fixed() +
  guides(color = guide_legend(override.aes = list(alpha = 1)))
```
<button onclick="myFunction(AfH2K7Wc0h)">Code</button>
<div id="AfH2K7Wc0h" style="display: none">
```{r eval=FALSE}
water_raw %>%
  filter(
    country_name == "Sierra Leone",
    lat_deg > 0, lat_deg < 15, lon_deg < 0,
    status_id %in% c("y", "n")
  ) %>%
  ggplot(aes(lon_deg, lat_deg, color = status_id)) +
  geom_point(alpha = 0.1) +
  coord_fixed() +
  guides(color = guide_legend(override.aes = list(alpha = 1)))
``` 
</div>
<button onclick="myFunction(kaDimFIfN7)">Explanation</button>
<div id="kaDimFIfN7" style="display: none">

</div>
</figure>



**To be continued...**
